Redshift Evolution of the Type Ia Supernova Population Stretch Distribution

### 2020-01-27 Nicolas.tex
Big changes. Version 1.5 at least!

\begin{document}
%\title{Evolution of the Type Ia Supernova Underlying Stretch Distribution as a Function of Redshift}
\title{Redshift Evolution of the Type Ia Supernova Population Stretch Distribution}

%\subtitle{II. An example text with infinitesimal scientific value whose title
%and subtitle may also be split}

@@

%\offprints{R. Plemmons, \email{plemmons@...}}

@@

\section{Introduction}

%\begin{figure*}
%    \centering
%    \subfigure[Method]{\label{fig:zdet}%        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}

%        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}
%    \subfigure[Result]{\label{fig:surveys_cuts}
%    \includegraphics[width=8.5cm]{A    \subfigure[Method]{\label{fig:zdet}
%rticle_figures/surveys_cuts_x.pdf}}
%    \subfigure[Method]{\label{fig:zdet}

%    \caption{(a) $z\St{max}$ determination from the magnitude limited data; (b) SDSS, %SNLS and PS1 samples cut at $z\St{max}$ as explained in
%    section \ref{sec:sam}. The data we use is in plain markers.}
%\subfigure[Method]{\label{fig:zdet}

%\end{figure*}

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate $z<0.5$, when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES
project][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
by cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM\nn{,} or when the SN luminosity is anchored at intermediate
redshift by the baryon acoustic oscillation (BAO) scale
\citep{riess2019,reid2019,planck2016, feeney2019}.  While using the tip of the
red giant branch technique in place of the Cepheids seem to favor an
\mr{intermediate} value of H$_0$ \citep{freedman2019}, time delay measurements
from strong lensing seem to also favor high H$_0$ values \citep{wong2019}.

@@

We present fig.~\ref{fig:maglim} the lightcurve stretch and color distributions
of SNe~Ia from these surveys, namely: PanStarrs \citep[PS1][]{rest2014}, the
Sloan Digital Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy
Survey \citep[SNLS][]{astier2006}. An ellipse in the salt2.4 $(x_1, c)$ plan
with semi-major axis $x_1=3$ and semi-minor axis $c=0.3$ encapsulates the full
distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
\citet{campbell2013} for similar contours, the second using a more conservative
$c=0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and  $c=0$ is
$M_0=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute
standardized magnitude at maximum of light along the aforementioned ellipse
given the standardization coefficient $\alpha=0.156$ and $\beta=3.14$ from
\cite{scolnic2018a}. The faintest SN~Ia is that with $x_1=-1.66$ and $c=0.25$
and has an absolute standardized magnitude at peak in Bessel-b band of
$-18.31\,\mathrm{mag}$. Since one need to detect this object typically a week
before and 10 days after peak to build a suitable lightcurve, the limiting
\nn{standardized} absolute magnitude effectively is approximately
$-18.00\,\mathrm{mag}$. Hence, given the $5\sigma$ point source detection
magnitude limit of a magnitude limited survey $m_{lim}$, one can derive the
maximum redshift above which the faintest SNe~Ia will be missed \nn{using the
link between apparent magnitude, redshift and absolute magnitude given by:}
\begin{equation}\label{eq:m_lim}
    \nn{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\,\text{mag}}
    < m_{lim}}
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_all.pdf}
    \caption{SALT2.4 stretch and color lightcurve parameters of SNe~Ia from the
        SDSS, PS1 and SNLS samples from the pantheon catalog. The individual SN
        data are shown as blue dots and a 2D histogram is shown in gray to
        highlight point density.  The ellipse ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$)
    is displayed, colored by the corresponding standardized absolute magnitude
using the $\alpha$ and $\beta$ coefficients from \citep{scolnic2018a}.}
    \label{fig:maglim}
\end{figure}

SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
filter that has a 24.8 mag $5\sigma$
depth\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{lim}=0.60$, in perfect
agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.
Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
agreement with, e.g., Figure~6 of \cite{scolnic2018a}. This figure could also
suggest a more conservative $z_{lim}$ of 0.27; both will be considered as
discussed below.  In similar redshift range, SDSS has a limiting magnitude of
22.5 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources.
\cite{kessler2009} \nn{pointed out} that during year-1 of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to the
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were used, such that \cite{kessler2009} and
\cite{dilday2008} show a relative completeness up to $z_{lim}=0.2$.  Following
these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit for
the rest of \nn{this study}.  For the rest of the analysis, we will consider
both cases, the \nn{aforementioned} redshift limits as well as more conservative
cuts for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$
for PS1 and $z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}.

The complete sample selection is summarized table~\ref{tab:sample} and the
redshift distribution of these three surveys in are shown fig.~\ref{fig:cuts}.
We see in this figure that the redshift limit we have selected roughly
correspond to the peak of these histograms as expected.

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). As the searches were much deeper than the spectrophotometric follow-up
made by SNfactory \nn{(SNf)}, SNe~Ia within a redshift range of $0.02<z<0.09$
\citep[as in ][]{rigault2018} should also be a random sampling of the underlying
SN population. The SNf data is particularly useful for studying SN property
drift as it enable us to have a large SN~Ia sample at $z<0.1$.  The HST sample
from Pantheon follows the same logic of having a search deeper than the
follow-up and we therefore kept it entirely \citep{FIND BACK THE REF}, see
table~\ref{tab:sample}.

\begin{table}
    \centering
    \caption{\nn{Source surveys and number of SNe~Ia used in our "complete"
    sample. Conservative cuts are indicated in brackets.}}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & -- & 114\\[0.30em]
        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
        HST & -- & 26\\[0.30em]
        Total & -- & 569(422)\\[0.30em]
        \hline
    \end{tabular}
\end{table}

@@

%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, fig. 10 suggest still
%okat z=0.2}

@@

In \cite{rigault2018} we presented a modeling of the evolution of the fraction
of prompt and delayed SNe~Ia as a function of redshift following former work on
rates and delay time distributions \citep[e.g.,][]{mannucci2005,
scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
In short, we assumed that the number of prompt SNe~Ia follows the star formation
activity in the Universe while the number of delayed SNe~Ia follows the number
of Gyr-old stars in the Universe, i.e. the stellar mass. Hence, if we denote
$\delta(z)$ (resp. $\psi(z)$) the fraction of young (resp. old) SNe~Ia in the
Universe as a function of redshift, then their ratio $\delta/\psi$ is expected
to follow the evolution of the specific star formation rate which goes as
$(1+z)^{2.8}$ until $z\sim2$ \citep[e.g.,][]{tasca2015}. Hence since
\nn{$\delta(0.05) \sim \psi(0.05)$} \citep{rigault2013,rigault2018,wiseman2020},
in agreement with rate expectations \citep{mannucci2006,rodney2014}, \nn{and
$\delta + \psi = 1$,} then we find in \cite{rigault2018}:
\begin{align}
    \label{eq:delta}
    \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
    \mathrm{and}\\
    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1}
\end{align}
where $K=0.87$. This modelisation is comparable to the evolution predicted by
\cite{childress2014} based on SN rates in galaxies depending on their quenching
time as a function of their stellar mass.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
    \caption{\textit{Main}: Salt2.4 lightcurve stretch ($x_1$) as a function of
        the local specific star formation rate (LsSFR) for SNf data used in this
        analysis. The color corresponds to the probability for the SNe~Ia to be
        above $\log(\mathrm{LsSFR}=-10.82)$, which defines $p(y)$; see
        \cite{rigault2018}. \textit{Right}: histogram of $p(y)$-weighted
        marginalization of the SN stretch.  On both panel, we illustrate the
        Base model fitted on the SNf data. The young and old population
        contributions are shown in purple and yellow, respectively.}
    \label{fig:stretchlssfr}
\end{figure*}

\nn{In \cite{rigault2018} we presented the relation between SNe stretches and
LsSFR measurements, which trace progenitor age, from the SNf sample, and is
shown fig.~\ref{fig:stretchlssfr}}. To model the evolution of the SN stretch as
a function of redshift given our aforementioned model of the evolution of the
fraction of young and old SNe~Ia with redshift, we need to model the SN stretch
distribution for each age subsample. Given the structure of the stretch--LsSFR
scatter plot shown in fig.~\ref{fig:stretchlssfr}, we define our model as
follows: for the young population the underlying stretch distribution is modeled
as $\mathcal{N}(\mu_1, \sigma_1^2)$, i.e. a normal distribution centered on
$\mu_1$ with a $\sigma_1$ width; the old population stretch distribution is
modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
\mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a gaussian mixture where one mode is the
same as the young population one.  \nn{When combining our prediction of the
evolution of young SNe~Ia as a function of redshift eq.~\ref{eq:delta}, and the
stretch distribution of both young and SNe~Ia, our model for the underlying
distribution of stretch $X_1\left(z\right)$ as a function of redshift is:}

\begin{align}
    \label{eq:stretchz}
    \nn{X_1\left(z \right) =
    }&\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
    (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
\end{align}

@@

\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
    \mathrm{d}x^{i}_{1}, p(\nn{\text{young}})^{i}}\right),
\end{equation}

\nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
$p(\text{young})^{i}$ are the salt2.4 stretch, error of the salt2.4 stretch and
the probability that the SN is young, respectively. Depending on the data
available, there are two ways to fit these 5 free parameters}.

@@

\nn{When using SNf data, we choose $p(\text{young})$ to be their given $p(y)$,
computed from their LsSFR data (see fig.~\ref{fig:stretchlssfr}). Thus, we would
have $\mathcal{P}$ from eq.~\ref{eq:likelihood} to be}

\begin{align}
    \label{eq:likelihoodsnf}
    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, p(\nn{\text{young}})^{i}} =
    p(y)^{i} \times &\mathcal{N}\left(\mu_1,
    \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
    \nonumber\\
    (1-p(y)^{i}) \times \nn{\bigg[} a \times & \mathcal{N}\left(\mu_1,
        \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
        \nonumber\\
     (1-a) \times &\mathcal{N}\left(\mu_2,
 \sqrt{\sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) \nn{\bigg]}
\end{align}

In practice, since "$a$" is defined between 0 and 1, we fit for $\alpha$ such
that $a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.
\mr{Results on fitting the SNf data with this model are shown
table~\ref{tab:modelresults} and illustrated in fig.~\ref{fig:stretchlssfr}.}

@@

Given eq.~\ref{eq:stretchz}, we can extend the analysis to non-SNfactory sample
by fitting the free parameters of the model
($\theta\equiv{\mu_1,\mu_2,\sigma_1,\sigma_2,a}$) assuming the evolution of the
fraction of young SNe~Ia as a function of redshift given by $\delta(z)$ and
given, this time, $x_1^{i}$,  $\mathrm{d}x^{i}_{1}$ and $z^{i}$, the stretch,
error on the stretch and redshift of any given SN $i$. In practice, it means
that we are still minimizing for the same eq.~\ref{eq:likelihood}, but replacing
$p(y)^{i}$ by $\delta(z^{i})$ eq.~\ref{eq:likelihoodsnf}.

\mr{For the rest of the analysis, we will therefore fit for
eq.~\ref{eq:likelihood} using $p(y)^{i}$ --~the probability for the SN $i$ to be
young~-- when available (i.e. for SNf dataset) and $\delta(z^{i})$ --~the
expected fraction of young SNe~Ia at the SN redshift $z^{i}$~-- otherwise.}

\mr{Results on fitting the all the 569(422) SNe~Ia data (with conservative cuts)
with this model are given table~\ref{tab:modelresults} and illustrated in
fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
stretch per bin of redshift follow our redshift drift modeling. That is, when
considering selection-bias free SN~Ia samples, SNe~Ia at higher redshift have on
average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
redshift ($-0.17\pm 0.10$ at $z\sim0.05$). This is indeed what is expected if
old environments favor low SN stretch \citep[e.g.]{howell2007} \textit{and} if
the fraction of old SNe~Ia reduces as a function of redshift. See
section~\ref{sec:results} for a more quantitative discussion.}

\begin{figure*}
    \centering
    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
    \caption{Evolution of the mean SN salt2.4 stretch ($x_1$) as a function of
        redshift.  Markers show the mean stretch measured in redshift bins of
        equal sample size indicated in light gray at the bottom of each redshift
        bins. Full and light markers are used when considering the nominal or
        the conservative sample, respectively, and are colored by their mean
        redshift.  The orange horizontal line represent the mean redshift of the
    nominal sample illustrating the expectation if the SN stretch distribution
is not drifting. Best fits of our Base drifting model are shown as blue,
dashed-blue and gray, when fitted on the nominal, the conservative or the SNf
dataset, respectively; all are similar. The light-blue illustrate the amplitude
of the error of the best fit model when considering the nominal dataset.}
    \label{fig:modelall}
\end{figure*}

@@

\mr{Considering or not the conservative sample has no effect on our conclusion
as illustrated fig.~\ref{fig:modelall} and presented table~\ref{tab:comp}.}

>>>>

\begin{document}
\title{Redshift Evolution of the Type Ia Supernova Population Stretch
Distribution}

@@

@@

\section{Introduction}
Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate $z<0.5$, when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES
project][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
by cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM\nn{,} or when the SN luminosity is anchored at intermediate
redshift by the baryon acoustic oscillation (BAO) scale
\citep{riess2019,reid2019,planck2016, feeney2019}.  While using the tip of the
red giant branch technique in place of the Cepheids seem to favor an
\mr{intermediate} value of H$_0$ \citep{freedman2019}, time delay measurements
from strong lensing seem to also favor high H$_0$ values \citep{wong2019}.

@@

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
distributions of SNe~Ia from these surveys, namely: PanStarrs
\citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
\citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plan with semi-major axis $x_1=3$ and semi-minor axis $c=0.3$ encapsulates
the full distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011}
and \citet{campbell2013} for similar contours, the second using a more
conservative $c=0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and
$c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute
standardized magnitude at maximum of light along the aforementioned ellipse
given the standardization coefficient $\alpha=0.156$ and $\beta=3.14$ from
\cite{scolnic2018a}. The faintest SN~Ia is that with $x_1=-1.66$ and $c=0.25$
and has an absolute standardized magnitude at peak in Bessel-B band of
$-18.31\,\mathrm{mag}$. Since one need to detect this object typically a week
before and 10 days after peak to build a suitable lightcurve, the limiting
\nn{standardized} absolute magnitude effectively is approximately
$-18.00\,\mathrm{mag}$. Hence, given the $5\sigma$ point source detection
magnitude limit of a magnitude limited survey $m_{lim}$, one can derive the
maximum redshift above which the faintest SNe~Ia will be missed \nn{using the
relation between apparent magnitude, redshift and absolute magnitude given by:}
\begin{equation}\label{eq:m_lim}
    \nn{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\,\text{mag}}
    < m_{lim}}
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch and color lightcurve parameters
        of SNe~Ia from the SDSS, PS1 and SNLS samples from the pantheon catalog.
        The individual SN data are shown as blue dots and a 2D histogram is
        shown in gray to highlight point density. The ellipse
        ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$) is displayed, colored by the
        corresponding standardized absolute magnitude using the $\alpha$ and
        $\beta$ coefficients from \citep{scolnic2018a}. \nn{The lines represent
        the $(x_1, c)$ evolution for $m = m_{lim}$ in eq.~\ref{eq:m_lim},
    for $z=0.60$ and $z=0.70$ using SNLS's $m_{lim}$ of $24.8\,$mag. The data
    above these lines are those for which $m > m_{lim}$, meaning they can't be
    observed.}}
    \label{fig:maglim}
\end{figure}

SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
filter that has a 24.8 mag $5\sigma$
depth\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{lim}=0.60$, in perfect
agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.
Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
agreement with, e.g., Figure~6 of \cite{scolnic2018a}. This figure could also
suggest a more conservative $z_{lim}$ of 0.27; both will be considered as
discussed below.  In similar redshift range, SDSS has a limiting magnitude of
22.5 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources.
\cite{kessler2009} \nn{pointed out} that during year-1 of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to the
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were used, such that \cite{kessler2009} and
\cite{dilday2008} show a relative completeness up to $z_{lim}=0.2$. Following
these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit for
the rest of \nn{this study}. For the rest of the analysis, we will consider both
cases, the \nn{aforementioned} redshift limits as well as more conservative cuts
for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for
PS1 and $z_{lim}=0.55$ for SNLS (following Fig.~14 of \citealt{perrett2010}).

The complete sample selection is summarized table~\ref{tab:sample} and the
redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
We see in this figure that the redshift limits we have selected roughly
correspond to the peak of these histograms\nn{. As the peaks indicate that we
are starting to miss the faintest SNe~Ia, it seems logical to have the redshift
cuts fall near said peaks.}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). \nn{Their spectroscopic follow-up were only done on a redshift range of
$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
deeper. As such, these SNe~Ia} should also be a random sampling of the underlying
SN population. The SNf data is particularly useful for studying SN property
drift as it enable us to have a large SN~Ia sample at $z<0.1$.  The HST sample
from Pantheon follows the same logic of having a search deeper than the
follow-up and we therefore kept it entirely \citep{FIND BACK THE REF}, see
table~\ref{tab:sample}.

\begin{table}
    \centering
    \caption{\nn{Source surveys and number of SNe~Ia used in our ''complete''
    sample. Conservative cuts are indicated in brackets.}}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & -- & 114\\[0.30em]
        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
        HST & -- & 26\\[0.30em]
        Total & -- & 569(422)\\[0.30em]
        \hline
    \end{tabular}
\end{table}

@@

%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, Fig. 10 suggest still
%okat z=0.2}

@@

In~\cite{rigault2018} we presented a modeling of the evolution of the fraction
of prompt and delayed SNe~Ia as a function of redshift following former work on
rates and delay time distributions \citep[e.g.,][]{mannucci2005,
scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
In short, we assumed that the number of prompt SNe~Ia follows the star formation
activity \nn{(SFR)} in the Universe while the number of delayed SNe~Ia follows
the number of Gyr-old stars in the Universe, i.e. the stellar mass \nn{(M)}.
Hence, if we denote $\delta(z)$ (resp. $\psi(z)$) the fraction of young (resp.
old) SNe~Ia in the Universe as a function of redshift, then their ratio
$\delta/\psi$ is expected to follow the evolution of the specific star formation
rate \nn{(SFR/M)} which goes as $(1+z)^{2.8}$ until $z\sim2$
\citep[e.g.,][]{tasca2015}. Hence since \nn{$\delta(0.05) \sim \psi(0.05)$}
\citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
\citep{mannucci2006,rodney2014}, \nn{and $\delta + \psi = 1$,} then we find in
\cite{rigault2018}:
\begin{align}
    \label{eq:delta}
    \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
    \mathrm{and}\\
    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1}
\end{align}
where $K=0.87$. This modelisation is comparable to the evolution predicted by
\cite{childress2014} based on SN rates in galaxies depending on their quenching
time as a function of their stellar mass.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$) as a function of
        the local specific star formation rate (LsSFR) for SNf data used in this
        analysis. The color corresponds to the probability for the SNe~Ia to be
        above $\log(\mathrm{LsSFR}=-10.82)$, which defines $p(y)$; see
        \cite{rigault2018}. \textit{Right}: histogram of $p(y)$-weighted
        marginalization of the SN stretch. On both panel, we illustrate the
        Base model fitted on the SNf data. The young and old population
        contributions are shown in purple and yellow, respectively.}
    \label{fig:stretchlssfr}
\end{figure*}

\nn{In~\cite{rigault2018} we presented the relation between SNe stretches and
LsSFR measurements, which trace progenitor age, from the SNf sample, and is
shown in Fig.~\ref{fig:stretchlssfr}}. To model the evolution of the SN stretch
as a function of redshift, given our aforementioned model of the evolution of
the fraction of young and old SNe~Ia with redshift, we need to model the SN
stretch distribution for each age subsample. Given the structure of the
stretch--LsSFR scatter plot shown in Fig.~\ref{fig:stretchlssfr}, we define our
model as follows: for the young population the underlying stretch distribution
is modeled as $\mathcal{N}(\mu_1, \sigma_1^2)$, i.e. a normal distribution
centered on $\mu_1$ with a $\sigma_1$ width; the old population stretch
distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
\mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a gaussian mixture where one mode is the
same as the young population one. \nn{When combining our prediction of the
evolution of young SNe~Ia as a function of redshift (eq.~\ref{eq:delta}),
and the stretch distribution of both young and old SNe~Ia, our model for the
underlying distribution of stretch $X_1\left(z\right)$ as a function of
redshift is:}
\begin{align}
    \label{eq:stretchz}
    \nn{X_1\left(z \right) =
    }&\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
    (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
\end{align}

@@

\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
    \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}}\right),
\end{equation}

\nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
    $p_y{}^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
    \textsc{\texttt{SALT2.4}} stretch and the probability that the SN is young,
    respectively. Depending on the data available, there are two ways to fit
these 5 free parameters}.

@@

\nn{When using SNf data, we choose $p_y{}^i$ to be their given
$p_y(\mathrm{LsSFR}^i)$, computed from their LsSFR data (see
Fig.~\ref{fig:stretchlssfr}). Thus, we would have $\mathcal{P}$ from
eq.~\ref{eq:likelihood} to be}

\begin{align}
    \label{eq:likelihoodsnf}
    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}} =
    \nn{p_y(\mathrm{LsSFR}^i)} \times &\mathcal{N}\left(\mu_1,
    \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
    \nonumber\\
    (1-\nn{p_y(\mathrm{LsSFR}^i)}) \times \nn{\bigg[} a \times &
        \mathcal{N}\left(\mu_1,
        \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
        \nonumber\\
     (1-a) \times &\mathcal{N}\left(\mu_2,
 \sqrt{\sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) \nn{\bigg]}
\end{align}

In practice, since ''$a$'' is defined between 0 and 1, we fit for $\alpha$ such
that $a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.
\mr{Results on fitting the SNf data with this model are shown
table~\ref{tab:modelresults} and illustrated in Fig.~\ref{fig:stretchlssfr}.}

@@

Given eq.~\ref{eq:stretchz}, we can extend the analysis to non-SNfactory sample
by fitting the free parameters of the model ($\theta\equiv{\mu_1, \mu_2,
\sigma_1, \sigma_2, a}$) assuming the evolution of the fraction of young SNe~Ia
as a function of redshift given by $\delta(z)$ and given, this time, $x_1^{i}$,
$\mathrm{d}x^{i}_{1}$ and $z^{i}$, the stretch, error on the stretch and
redshift of any given SN $i$. In practice, it means that we are still minimizing
for the same eq.~\ref{eq:likelihood}, but \nn{using $p_y{}^i = \delta(z^{i})$,
which is equivalent to replacing $p_y(\mathrm{LsSFR}^i)$ with $\delta(z^i)$ in
eq.~\ref{eq:likelihoodsnf}.}

\mr{For the rest of the analysis, we will therefore fit for
eq.~\ref{eq:likelihood} using \nn{$p_y(\mathrm{LsSFR}^i)$} --~the
probability for the SN $i$ to be young~-- when available (i.e. for SNf
dataset) and $\delta(z^{i})$ --~the expected fraction of young SNe~Ia at the
SN redshift $z^{i}$~-- otherwise.}

\mr{Results on fitting the all the 569(422) SNe~Ia data (with conservative cuts)
with this model are given table~\ref{tab:modelresults} and illustrated in
Fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
stretch per bin of redshift follow our redshift drift modeling. That is, when
considering selection-bias free SN~Ia samples, SNe~Ia at higher redshift have on
average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
redshift ($-0.17\pm 0.10$ at $z\sim0.05$). This is indeed what is expected if
old environments favor low SN stretch \citep[e.g.]{howell2007} \textit{and} if
the fraction of old SNe~Ia reduces as a function of redshift. See
section~\ref{sec:results} for a more quantitative discussion.}

\begin{figure*}
    \centering
    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
    \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        as a function of redshift.  Markers show the mean stretch measured in
        redshift bins of equal sample size indicated in light gray at the bottom
        of each redshift bins. Full and light markers are used when considering
        the nominal or the conservative sample, respectively, and are colored by
        their mean redshift.  The orange horizontal line represent the mean
        redshift of the nominal sample illustrating the expectation if the SN
        stretch distribution is not drifting. Best fits of our Base drifting
        model are shown as blue, dashed-blue and gray, when fitted on the
    nominal, the conservative or the SNf dataset, respectively; all are similar.
The light-blue illustrate the amplitude of the error of the best fit model when
considering the nominal dataset.}
    \label{fig:modelall}
\end{figure*}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/bbc_comp_PS1.pdf}
    \caption{Underlying SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        distributions for the PS1 sample. The Green line(band) shows our best
        fitted asymmetric Gaussian distribution (its error) mimicking the BBC
        approach. The dashed line shows that estimated by \cite{scolnic2018a}.
    The blue to red curves display our Base drifting stretch model estimated at
each PS SN redshift; bluer for lower redshifts.}
    \label{fig:bbc_pdf_ps1}
\end{figure}
